---
title: Quantified Grad Student Analysis
ouput: html_document

---


```{r, echo=FALSE}
# setwd("~/Dropbox/proj/qgs/")
```

```{r setup, message=FALSE}
knitr::opts_chunk$set(tidy = FALSE, fig.height=4, cache=TRUE)
options(stringsAsFactors=FALSE)
## Loading Packages
require(dplyr)
require(lubridate)
require(ggplot2)
```

## Outline
I'll be working with two datasets:

- `clock`: which is when I come into campus, and when I leave
- `gleeo`: which is the tracked time I spent on different tasks throughout each day

In this analysis I am going to:

1. load in the data
 - add in a week count to gleeo and clock
 - add the duration of campustime in clock

2. Answer the question: *How much time do I invest in Grad School?*, which can be broken down into
 - how much time do I spend on campus, per day / week  
 - how much time am I working on grad school stuff, per day / week  
 - What percent of time that I am on campus am I actually getting work done?
  - How efficient am I? (account for time worked not at campus)
 - At the end I want to see "How many hours do I work a week", so that has to be one output.
 
 
 


## import the data
```{r}
gleeo.df <- read.csv(file="2014-05-31_GleeoExport.csv", dec=".")
clock.df <- read.csv(file="2014-06-15_clocktime.csv")
# Renaming gleeo.df$Start.Date as date for ease of use
names(gleeo.df)[4] <- "date"

```


## Modify the data
### adding week count to `gleeo.df` and `clock.df`
I want to add in a week count, weeks start on Sunday. The reason for this is that I tend to think in week long work blocks.

```{r}
# saturday
format(as.Date("2014-06-7"), "%U")
week("2014-06-07")
# sunday
format(as.Date("2014-06-8"), "%U")
```
So, apparently I can set something as a date, and then use the format function to get the week number. It seems that the week starts by default on a Sunday, so that's neat.

since my times don't span over a day, I only need to worry about one of the dates in my `gleeo` data frame

```{r}
# non-lubridate method
# gleeo.df$week <- format(ymd(gleeo.df$date), "%U")
#clock.df$week <- format(ymd(clock.df$date), "%U")

gleeo.df$week <- week(gleeo.df$date)
clock.df$week <- week(clock.df$date)
```

> I had the date in clocktime set as 3/3/2014, which is not the same as 2014-03-03, and hence `as.Date()` had issues with it. So I changed it in my google docs. Also it's nice to have clock and gleeo have the same date format.

### Calculating time stuff
Go Go Lubridate

### Adding the weekday to `gleeo.df` and `clock.df`

```{r}
gleeo.df$day <- wday(gleeo.df$date, label=TRUE)
clock.df$day <- wday(clock.df$date, label=TRUE)
```


### extra days from the `clock.df`
Turns out I had an extra week in the beginning, and then a couple days at the end. So there's more days in clock.df than there is in gleeo.df

```{r}
# clock.df <- clock.df[ymd(clock.df$date) %in% ymd(gleeo.df$date), ]
```

### Getting Tracked time

Here I'm just going to generate a lot of values, and deal with organizing them at the end.

```{r}
# Calculating campustime
clock.df$campustime <- interval(
         ymd_hms(paste(clock.df$date, clock.df$in_time)),
         ymd_hms(paste(clock.df$date, clock.df$out_time))
         ) / ehours(1)
```

## How much time do I invest in Grad School?


### how much time do I spend on campus, per day / week

Overall, what is the amount of time I spend on campus per day, let's look at this with a density plot:
```{r}
ggplot(clock.df, aes(x=campustime)) + 
    geom_histogram(aes(y=..density..),
                   binwidth=.25, color="black", fill="white") + 
    geom_density(alpha=.3, fill="black")


```



Thinking about this, I expect there to be some structure to the times here, as I take a certain bus in, and then take a few possible busses back, so that would explain the lumps in the distribution.
```{r}
clock.df[which(clock.df$campustime <=5), ]
# row 47, 2014-05-16
# clock.df<- filter(clock.df, date!= "2014-05-16" )
```

The outlier on Friday is one day when I left early, I was there only a half day (prelim season, just wanted to leave early).

```{r}
summary(clock.df$campustime)
```

So, my mean time spent on campus is `r summary(clock.df$campustime)[4]` hrs per day.


For how much time I spend on campus per day, it might be interesting to look at the day (Monday - Sunday)

```{r}
clock.df %>% group_by(day) %>% summarize(daycampus = mean(campustime))
clock.df %>% group_by(day) %>% summarize(daycampus = sd(campustime))

anova(lm(campustime ~ day, data=clock.df))
# marginally signifacnt effect of day on the amount of time I spend on campus, surprising.

ggplot(clock.df, aes(x=day, y=campustime)) + geom_violin()
```

From the plots there's definitely some outliers on Tues and Wed potentially, and there's the half day on Friday coming in at around 4hrs.   


#### and now, per week

Let's look at the distribution of the data:
```{r}
weekclock <- clock.df %>% group_by(week) %>% summarize(weekcampus = sum(campustime))
# is there a week effect?
anova(lm(campustime ~ week, data=clock.df))
# nope

ggplot(weekclock, aes(x=weekcampus)) + 
    geom_histogram(aes(y=..density..),
                   binwidth=1, color="black", fill="white") + 
    geom_density(alpha=.3, fill="black")

```
I guess I really don't have enough points to see an interesting distribution. Maybe there's a bimodal distribution. When I get more data I can see if that is actualy the case (maybe I have productive weeks, and unproductive weeks, and no inbetween) 

Ok, what about a summary of the data
```{r}
summary(weekclock$weekcampus)
```

so, on average I spend `r summary(weekclock$weekcampus)[4]` hours on campus a week.

#### In summary
I spend a mean of `r summary(clock.df$campustime)[4]` hrs per day, and `r summary(weekclock$weekcampus)[4]` hrs per week at campus





### how much time am I working on grad school stuff, per day / week
  - this is the gleeo tracked time
  
```{r}
# Adding tracked time per day (trackedtime)
gleeotime <- gleeo.df %>% group_by(date) %>% 
    summarize(tottrackedtime = sum(Decimal.Duration))

## adding time corrected for personal time
projtimes <- gleeo.df %>% group_by(Project, date) %>%
    summarize(totime = sum(Decimal.Duration))


perstime <- projtimes[which(projtimes$Project == "personal"), 2:3]
names(perstime)[2] <- "perstime"

# left join keeps all the rows of x, not y
trackedtime <- left_join(x=gleeotime, y=perstime, by="date")
# missing values of perstime are NAs, need to be zeros for further work.
trackedtime$perstime[is.na(trackedtime$perstime)] <- 0

# prodtime, tracked time doing productive things
trackedtime$prodtime <- trackedtime$tottrackedtime - trackedtime$perstime

trackedtime$day <- wday(ymd(trackedtime$date), label=TRUE)
trackedtime$week <- format(ymd(trackedtime$date), "%U")
```

Overall, how much productive time do I have per day?
```{r}
ggplot(trackedtime, aes(x=prodtime)) +
    geom_histogram(aes(y=..density..),
                   binwidth=.25, color="black", fill="white") + 
    geom_density(alpha=.3, fill="black")
```
Well, looks like my most common productive time is around 5 hours. The only thing is that this includes Saturday and Sunday. Let's remove those and look at what I have

```{r}
trackedtime_MF <- trackedtime %>% filter(day != "Sat", day != "Sun")
summary(trackedtime_MF$prodtime)
summary(trackedtime$prodtime)
# mean is definitely different
ggplot(trackedtime_MF, aes(x=prodtime)) +
    geom_histogram(aes(y=..density..),
                   binwidth=.25, color="black", fill="white") + 
    geom_density(alpha=.3, fill="black")
```

So the mean productive time per work day is `r summary(trackedtime_MF$prodtime)[4]` hours.

For further work I am going to only be looking at weekday prodtime



How does my prodtime look as a function of each day?
```{r}
anova(lm(prodtime ~ day, trackedtime_MF))
ggplot(trackedtime, aes(x=day, y=prodtime)) + 
    geom_violin()
```

Is my prodtime a function of week?
```{r}
anova(lm(prodtime ~ week, trackedtime_MF))
ggplot(trackedtime, aes(x=week, y=prodtime)) + 
    geom_violin()
ggplot(trackedtime_MF, aes(x=week, y=prodtime)) + 
    geom_violin()
ggplot(trackedtime_MF, aes(x=week, y=prodtime)) + 
    geom_line()
```


How much prodtime do I have per week?

```{r}
trackedtime_MF_week <- trackedtime_MF %>% group_by(week) %>% summarize(weekprodtime = sum(prodtime))
summary(trackedtime_MF_week$weekprodtime)
ggplot(trackedtime_MF_week, aes(x=weekprodtime)) +
    geom_histogram(aes(y=..density..),
                   binwidth=.25, color="black", fill="white") + 
    geom_density(alpha=.3, fill="black")

binaryweek <- as.numeric(trackedtime_MF$week) %% 2

ggplot(trackedtime_MF, aes(x=date, y=prodtime, group=1, color=as.factor(binaryweek) )) + geom_line()
ggplot(trackedtime_MF, aes(x=date, y=prodtime, group=week, color=week) ) + geom_line()
ggplot(trackedtime_MF_week, aes(x=week, y=weekprodtime, group=1)) + geom_line()
```

Well, that distribution looks pretty crappy. More data will probably fill it in. 


#### Summary

My mean productive time per work day is `r summary(trackedtime_MF$prodtime)[4]` hours per day, and `r summary(trackedtime_MF_week$weekprodtime)[4]` hours per week.




### What percent of time that I am on campus am I actually getting work done?
For this section I am going to calculate the amount of time that I am actually working while I am at campus. This is the amount of tracked time that falls within the interval when I am on campus. Then I am subtracting the amount of personal time per day. 
```{r}
# pairing down clock.df into only those dates with tracked times
timeday.df <- clock.df[ ymd(clock.df$date) %in% ymd(gleeo.df$date), ]

campusinterval <- data.frame(date = timeday.df$date,
                             inter = interval(
         ymd_hms(paste(timeday.df$date, timeday.df$in_time)),
         ymd_hms(paste(timeday.df$date, timeday.df$out_time))
         ))

# pair down the intervals so there is only those dates with tracked times
campusinterval <- campusinterval[ymd(campusinterval$date) %in% ymd(gleeo.df$date), ]

gleeo.inst <- data.frame(date= gleeo.df$date,
                         inst= ymd_hm(paste(
                             gleeo.df$date, gleeo.df$Start.Time)))

# making a df that holds if the rows are in a campus interval
rowBoolean.df <- data.frame(date= gleeo.inst$date)
for(i in 1:length(campusinterval$inter)){
    rowBoolean.df<- cbind(rowBoolean.df, gleeo.inst$inst %within% campusinterval$inter[i])
}

# populate a new df with only the rows that are in the campus interval
    # grabbing the first iteration to make a df to catch the output
cinterGleeo.df <- gleeo.df[rowBoolean.df[, 2], ]
for(i in 3:length(names(rowBoolean.df))){
   cinterGleeo.df <- rbind(cinterGleeo.df,
                           gleeo.df[rowBoolean.df[, i], ])     
}

# calc tracked time using the new df
catch <- cinterGleeo.df %>% group_by(date) %>% summarize(trackedCampusInterval= sum(Decimal.Duration))

timeday.df$trackedCampusInterval <- catch$trackedCampusInterval

# let's add columns for the total tracked time, perstime, and prodtime from the 
# trackedtime data.frame
timeday.df <- left_join(timeday.df, trackedtime[1:4] )



#  calculate efficiency, prodtime
timeday.df$campusprodtime <- timeday.df$trackedCampusInterval-timeday.df$perstime
timeday.df$efficiency <- (timeday.df$campusprodtime) / timeday.df$campustime
```

Didn't use dplyr, but at least I got some values. Time to look at the distribution
### Distribution of Efficiency
```{r}
ggplot(timeday.df, aes(x=efficiency)) +
    geom_histogram(aes(y=..density..),
                   binwidth=.025, color="black", fill="white") + 
    geom_density(alpha=.3, fill="black")
summary(timeday.df$efficiency)[c(1,4,6)]
```



### Efficiency per day, M-F
```{r}
binaryweek <- as.numeric(timeday.df$week) %% 2

ggplot(timeday.df, aes(x=date, y=efficiency, group=1, color=as.factor(binaryweek) )) + geom_line()
ggplot(timeday.df, aes(x=date, y=efficiency, group=week, color=week )) +
    geom_line(size=2) + geom_point(color="black") + theme_bw()
ggplot(timeday.df, aes(x=date, y=efficiency)) + geom_point()
```


```{r}
anova(lm(efficiency ~ day, timeday.df))
ggplot(timeday.df, aes(x=day, y=efficiency)) + 
    geom_violin()

ggplot(timeday.df, aes(x=day, y=efficiency, group= week, color= week)) + geom_line()
```

With that last graph I was trying to figure out a visualization for looking at if there is a trend as the week goes on

### Efficiency per day, by week
```{r}
anova(lm(efficiency ~ week, timeday.df))
ggplot(timeday.df, aes(x=week, y=efficiency, group=1)) + geom_point()
```

So there is no week effect according to a linear model

```{r}
timeweek.df <- timeday.df %>% group_by(week) %>% summarize(aweff= mean(efficiency), totwtrackedCampusInterval = sum(timeday.df$campusprodtime))
ggplot(timeweek.df, aes(x=week, y=aweff, group=1)) + geom_line()
ggplot(timeweek.df, aes(x=week, y=totwtrackedCampusInterval, group=1)) + geom_line()
```

With that last graph I was trying to figure out a visualization for looking at how the trend changes per week


#### In Summary
The average amount of time that I work during the time I am at campus is `r summary((timeday.df$campusprodtime))[4]` hours per day, and `r summary(timeweek.df$totwtrackedCampusInterval)[4]` hours per week

## Relationships

### Productive time in the day for how much time I am at campus
This is looking at if the amount of stuff I get done during the day is affected by the amount of time I am on campus. Pretty much this is going to be looking at if I am at campus for a really long time, does the amount of work that I actually get done taper off.

```{r}
ggplot(timeday.df, aes(x=campustime, y=prodtime)) + geom_point()
anova(lm(prodtime ~ campustime, data=timeday.df))
```

### Productive time at campus for how much time I am at campus

```{r}
ggplot(timeday.df, aes(x=campustime, y=campusprodtime)) + geom_point()
anova(lm(campusprodtime ~ campustime, data=timeday.df))
```

### Efficiency for how much time I am at campus
```{r}
ggplot(timeday.df, aes(x=campustime, y=efficiency)) + geom_point()
anova(lm(efficiency ~ campustime, data=timeday.df))
```



### Amount of productive time given personal time?
```{r}
ggplot(timeday.df, aes(x=perstime, y=campusprodtime)) + geom_point()
anova(lm(campusprodtime ~ perstime, data=timeday.df))
```

## In summary
- I spend a mean of `r summary(clock.df$campustime)[4]` hrs per day, and `r summary(weekclock$weekcampus)[4]` hrs per week at campus
- My mean productive time per work day is `r summary(trackedtime_MF$prodtime)[4]` hours per day, and `r summary(trackedtime_MF_week$weekprodtime)[4]` hours per week. Only considering my at campus times 


## Saving useful Dataframes
This makes work in other report a little more concise
```{r}
save(clock.df, file="rep1_clock.df.Rdata")
save(gleeo.df, file="rep1_gleeo.df.Rdata")
save(timeday.df, file="rep1_timeday.df.Rdata")
save(cinterGleeo.df, file="rep1_cinterGleeo.df.Rdata")
save(trackedtime, file="rep1_trackedtime.df.Rdata")
```



