---
title: "Quantified Grad Student"
output: beamer_presentation
author: "Derek Nedveck"
date: "2014-06-24"
---

```{r, echo=FALSE}
knitr::opts_chunk$set(tidy = FALSE, cache=TRUE)
```


# What I'm going to present
- Intro - why did I do this
- data collection - what kind of data did I gather
- analysis
- future directions - other questions?


# Motivation
- entering grad school - how much time do I need to invest
- anecdotal: it will take all of your time

But, who am I?
========================================
Grad student in Plant Biology

- Genomics (computational work), and local adaptation (field / greenhouse / lab work)
- How do I compare to other Grad students?
 - I'm probably not very representative... but then what is a "normal" grad student?


Questions
======================
- How much time do I invest in Grad School?
 - how much time do I spend on campus?
 - how much time am I working on grad school stuff? 
 - What percent of time that I am on campus am I actually getting work done?


# Data Collection

## Tools

- Android App: Gleeo
- Google Spreadsheet on my phone

## Data collected

- Time doing work
- Time I'm on campus




# But how do I split my time up?

How many categories?

- for a week, I tracked everything in a spreadsheet to get a sense of what I was doing
- then, from that I tried to come up with categories to partition the tasks in the most meaningful way, with the least amount going into a "misc" bin
- I'm still thinking about how best to partition my time, and am likely going to implement another scheme for the fall semester (starting September)




Categories, cont. 
=================

 
 
# Analysis
```{r, echo= FALSE, message= FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
# go get data that I prepared in report1.Rmd, so I can load stuff already processed.

setwd("~/Dropbox/proj/qgs/")
load(file = "rep1_clock.df.Rdata")
load(file = "rep1_gleeo.df.Rdata")
load(file= "rep1_trackedTimePerCampusInterval.df.Rdata")
```

# How long am I at campus?
```{r, echo=FALSE, fig.height=6}
ggplot(clock.df, aes(x=campustime)) + 
    geom_histogram(aes(y=..density..),
                   binwidth=.25, color="black", fill="white") + 
    geom_density(alpha=.3, fill="black") + 
    ggtitle("Amount of time I spend at campus per day")
summary(clock.df$campustime)[c(1,4,6)]
```


# Grouped by weekday?

```{r, echo=FALSE, fig.height=6}
ggplot(clock.df, aes(x=day, y=campustime)) + geom_violin( fill = "dark gray")
```


# Time per week?
```{r, echo=FALSE, fig.height=6}
weekclock <- clock.df %>% group_by(week) %>% summarize(weekcampus = sum(campustime))

ggplot(weekclock, aes(x=weekcampus)) + 
    geom_histogram(aes(y=..density..),
                   binwidth=1, color="black", fill="white") + 
    geom_density(alpha=.3, fill="black")
summary(weekclock$weekcampus)[c(1,4,6)]
```

# How efficient am I with the time that I am at campus?

# In Summary

I spend a mean of **`r summary(clock.df$campustime)[4]` hrs per day**, and **`r summary(weekclock$weekcampus)[4]` hrs per week** at campus

Day:
```{r, echo=FALSE}
summary(clock.df$campustime)[c(1,4,6)]
```

Week:
```{r, echo=FALSE}
summary(weekclock$weekcampus)[c(1,4,6)]
```


Future work
============================
More questions, more data -- which are worth my time in pursuing?
- Rescuetime seems easy to implement and yield more data on what I am doing when I am working on a computer
- How much do I read?
 - time tracking data gets at this, but how many articles do I read?
 - I store my notes in markdown files, but this misses the ones that I read for my prelim (I read them, incorporated and cited them, without taking down notes)
 
- I want to get more "Deep Thought" time in (a la Cal Newport), what kind of metrics can I measure to view this
- what kind of metrics reflect success in Grad School?
- I've measured time in grad schol to get a sense of how much I am investing, but what can I measure to actually improve what I am doing?
 
 
Want more R?
==============================

TCRUG - Twin Cities R User Group meeting this Thursday, at 6:30pm on East Bank UMn
